---
title: "Creating Documents with R - October 2022"
subtitle: "Kevin O'Brien"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(knitr)
library(dplyr)
library(janitor)


```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```



### Step 1

* The relevant datasets, such as Harvest Block Details, Screening Information, is held on a master spreadsheet in a dedicated folder on Citrix.

* This Dataset is maintained by Emma Benson, John Landy, Marie Therese Roche and more.

* Citrix is the optimal location because of ease of access for the R programming environment,

---


### Step 2

* This data can be processed by the R programming language.

* In the first instance, the entire data set is loaded into the R environment using the ***{readxl}*** R package.

* Individual sheets from the master data spreadsheet are loaded as required.

<pre><code>
library(readxl)

HB_details <- read_excel("MasterData.xlsx", sheet = "HB_Details")

Screening <- read_excel("MasterData.xlsx", sheet = "Screening")
</code></pre>
---

### Step 3

* The data is reduced to the particular felling licence applications required for the current run.

* There are multiple types of reporting requirements:

  - Single Screened In Application - NIS and Prescreening Report
  - Multiple FLs applied for jointly - NIS and Prescreening Reports
  - Single Screened out Applications - Prescreening Report Only

* The relevant type of report required is also recorded.


* The data undergoes some pre-processing.



---

### Step 4

* One by one, data for each FL (or group of FLs) is extracted and processed by the main R programme.

* Using the ***officer*** R package, a word document is created and populated with information relevant to the each FL, along with formatted text.

* The ***officer*** R package can insert headings, images, tables and inserted sheets as appropriate. It can also specify landscape or portrait formats as required.

* Tables can be created using the ***flextable*** R package.




---


```{r child="officer.Rmd"}
```
```{r echo=FALSE}
### OFFICER
```

---


```{r child="flextable.Rmd"}
```
